### 2013-06-08 doc/pgas13/flat-combining.tex
Worked on FC framework section (w/ code examples)

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
class GlobalStack {
  GlobalAddress<T> top;
  Grappa::Mutex lock;
};
void push(GlobalAddress<GlobalStack> stack, T val) {
  // perform operation on the stack's master core
  delegate::call(stack.core(), [stack,val]{
    // force other operations to wait
    stack->lock.acquire();
    // do remote write to the top of the stack
    // (so other workers may interleave)
    delegate::write(stack->top, val);
    // bump the top pointer
    stack->top++;
    stack->lock.release();
  }); // blocks until response arrives
}
\end{lstlisting}

@@

% As discussed before, Grappa's distributed combining takes a different approach to expressing how operation locally  flat combining design takes a different approach to performing 
The Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine. As mentioned before, the local ``publication list'' is trivial to access atomically in Grappa due to workers being cooperatively scheduled.
Instead, in the Grappa FC Framework, a \emph{proxy} object is instantiated ahead of time on each core and each worker merges its update into that structure before blocking or becoming the combiner.
% In the original work, an explicit publication list is built up, and when the combiner scans over the list, it has a structure that it uses to accumulate requests in a logical way. In the case of the stack, for instance, it is a thread-local stack, which it can use to immediately match up pushes and pops. In the Grappa implementation, each worker's actions are atomic anyway, so rather than creating an explicit request list, the combiner object (or \emph{proxy}) is instantiated ahead of time for the data structure, and each worker does its own local combining into that structure before blocking. This reduces the space blowout of a large dynamically-sized publication list (with associated caching issues), and does not overly affect the expression of combining.
In this style, each global data structure must define the following:
\begin{enumerate}
  \item A local \emph{proxy} data structure for tracking updates to the global structure.
  \item \emph{Combining methods} that operate on the local proxy.
  \item A \emph{sync} method that globally commits the proxy's state.
\end{enumerate}

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
struct StackProxy {
  T values[CAPACITY];
  T* results[CAPACITY];
  ConditionVariable waitlist;
  int npush, npop;
  void push(T e) {
    if (npop > 0) { // satisfy locally
      *results[--npop] = e;
      waitlist.signal();
    } else {
      values[npush++] = e;
      block(waitlist);
    }
  }
  T pop() {
    T val;
    if (npush > 0) { // satisfy locally
      val = values[--npush];
      waitlist.signal();
    } else {
      pops[npop++] = &val;
      block(waitlist);
    }
  }
  void sync() {
    if (npush > 0) {

    } else if (npop > 0) {

      for (int i = 0; i < npop; i++) *results[i] = values[i];
    }
  }
};
\end{lstlisting}
\end{figure}

Being conservative about SC, workers typically block when issuing these requests, so the framework must also take care of delivering results and waking the correct workers. The flat-combining framework is responsible for ensuring that all of the combined operations eventually get committed. There are a number of ways progress could be guaranteed, but one of the simplest is to take an approach in the spirit of the original flat-combining, which is to ensure that as long as there are any outstanding requests, there is at least one worker committing a combined operation. When that combined operation finishes, if there are still outstanding requests that have built up in the meantime, another blocked worker is chosen to do another combined synchronization.

@@

The flat combining framework transparently wraps the \emph{proxy} object. Behind this layer of indirection, it manages multiple copies of the proxy so that, as discussed above, there is always at least one being committed and one to receive additional combining operations.

@@

\subsection{Global Stack}
To demonstrate the flat-combining framework, we revisit the Global Stack example from earlier. Recall that the ``master'' core owns the \texttt{top} index, which indicates where the top of the stack currently is in the stack's allocated global storage. In the non-combining case, each call to \texttt{push} must do an atomic increment of \texttt{top} to claim storage, then write to that location, then finally return.

>>>>

@@

% As discussed before, Grappa's distributed combining takes a different approach to expressing how operation locally  flat combining design takes a different approach to performing 
The Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine. As mentioned before, the local ``publication list'' is trivial to access atomically in Grappa due to workers being cooperatively scheduled.
Instead, in the Grappa FC Framework, a \emph{proxy} object is instantiated ahead of time on each core and each worker merges its update into that structure before blocking or becoming the combiner.
% In the original work, an explicit publication list is built up, and when the combiner scans over the list, it has a structure that it uses to accumulate requests in a logical way. In the case of the stack, for instance, it is a thread-local stack, which it can use to immediately match up pushes and pops. In the Grappa implementation, each worker's actions are atomic anyway, so rather than creating an explicit request list, the combiner object (or \emph{proxy}) is instantiated ahead of time for the data structure, and each worker does its own local combining into that structure before blocking. This reduces the space blowout of a large dynamically-sized publication list (with associated caching issues), and does not overly affect the expression of combining.
In this style, each global data structure must define the following:
\begin{enumerate}
  \item A local \emph{proxy} data structure for tracking updates to the global structure.
  \item \emph{Combining methods} that operate on the local proxy.
  \item A \emph{sync} method that globally commits the proxy's state.
\end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:stack_proxy}.

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
class GlobalStack {
  GlobalAddress<T> top;
  Grappa::Mutex lock;
};

void push(GlobalAddress<GlobalStack> stack, T val) {
  // perform operation on the stack's master core
  delegate::call(stack.core(), [stack,val]{
    // force other operations to wait
    stack->lock.acquire();
    // do remote write to the top of the stack
    // (so other workers may interleave)
    delegate::write(stack->top, val);
    // bump the top pointer
    stack->top++;
    stack->lock.release();
  }); // blocks until response arrives
}
\end{lstlisting}

\begin{figure}[t]
\centering
\begin{lstlisting}[style=grappa]
class GlobalStackProxy : Grappa::FCProxy {
  // pointer to data structure master
  GlobalAddress<GlobalStack> master;

  // local state for tracking pushes and pops
  T  pushed_values[1024];
  T* popped_results[1024];
  int npush, npop;

  // Combining Methods 
  // (either match locally or block for results)
  void push(T val);
  T pop();

  // Global sync (called by FC framework)
  void sync() override {
    if (npush > 0) {
      // on master, acquire lock and return GlobalAddress to the top of the stack
      auto top = delegate::call(master.core(),[=]{
        master->lock.acquire();
        return master->top;
      });
      // copy values to top of stack
      Grappa::memcpy(top, pushed_values, npush);
      // release lock at the master core
      delegate::call(master.core(),[master,npush]{
        master->top += npush;
        master->lock.release();
      });
    } else if (npop > 0) {} //...elided for space
  }
};
\end{lstlisting}
\caption{Snippet of the Proxy object for the GlobalStack. The code to synchronize pushes should resemble that from Figure~\ref{fig:push} but for multiple elements.}
\label{fig:proxy}
\end{figure}

With this data structure-specific proxy object, the FC framework does the bookkeeping to block and wake Workers, deliver results, manage when to synchronize, and choose the combiner.
The FC framework is responsible for ensuring that all of the combined operations eventually get committed. There are a number of ways progress could be guaranteed, but one of the simplest is to ensure that as long as there are any outstanding requests, at least one worker is committing a combined operation.
When that combined operation finishes, if there are still outstanding requests that have built up in the meantime, another blocked worker is chosen to do another combined synchronization.

@@

While a combining operation is in flight, new requests are likely to continue to accumulate. The framework transparently wraps the proxy object so that when it starts a sync on one, it can direct requests to a fresh combiner. It is important to note that combiners must start out fresh each time---it would be unsound for them to ``remember'' the previous updates and attempt to combine locally with them.

% The flat combining framework transparently wraps the \emph{proxy} object. Behind this layer of indirection, it manages multiple copies of the proxy so that, as discussed above, there is always at least one being committed and one to receive additional combining operations.

@@

\subsection{Global Stack and Queue}
Figure~\ref{fig:proxy} shows an excerpt from the definition of the proxy object for the Grappa GlobalStack. The proxy tracks pushes locally in the \texttt{pushed\_values} array. When \texttt{pop()} is called, if there are pushes in the local array, it immediately takes the value and wakes the last worker to push. Otherwise, it adds a pointer to a location on its stack where the \texttt{sync} operation, will eventually write the result. Because of local matching, when a Stack proxy is synchronized, it will have either all pops or all pushes, which makes the implementation of sync straightforward. Note that the operation to synchronize a batch of pushes looks almost identical to the code to do a single push from Figure~\ref{fig:push}.

The GlobalQueue has nearly the same implementation as the stack, but is unable to match locally, so \texttt{sync} first does the pushes, then does the pops.

\subsection{Global Set, and Map}
The Grappa GlobalSet uses a simple chaining design, implemented with a global array of cells (allocated from the global heap), which are partitioned evenly among all the cores, and indexed with a globally-agreed hash function. Our flat-combining version supports both \texttt{insert} and \texttt{lookup}. To track all of the keys waiting to be inserted, we use the hash set implementation from the C++11 standard library (\texttt{std::unordered\_set}). Similar to the case with pops, for lookups, we must keep track of pointers for where to put the results of lookup operations, which the proxy does with a hash map (again from the C++ standard library) from keys to a list of pointers. As discussed in Section~\ref{sec:sc}, matching lookups with inserts locally would force a particular sequential order, so instead we only allow matching inserts with inserts and lookups with lookups. This allows \texttt{sync} to simply issue all of the inserts and lookups in parallel and block until all have completed.

Our implementation of the GlobalMap matches that of the Set, but of course stores values of a given type. This does not change the requirements of the proxy except that the return values of lookups must be a copy of the value rather than just a boolean.
